\documentclass{article}
\usepackage{amssymb,latexsym,amsmath}

\title {WikiMathDB Idea Description}
\author{Dzmitry Lazerka}
\date{2007-11-01}


\begin{document}

\maketitle

\section{Introduction}

Idea of the WikiMathDB project was born while trying to understand a
mathematical proof while lot of theories were involved and lot of
"obvious" facts were omitted. Then I imagined: was there any way to
store the proof so detailed, that even computer can understand it?
The more I thought about it, the more I convinced of technical
possibility to build a computer system which goal is to contain
proofs in very detailed form; so detailed, that even computer can
check it. So, everyone can view at the proof in more detail. And
allow everyone to add his own theories. And provide different views
for various languages and terminologies. So, the idea was born.\\

Who needs in the WikiMathDB system? What the purpose of it?

At first, for educational purposes: if some student cannot
understand something "obvious" or forget a well-known thing, he
always can look in the WikiMathDB for explanation. Also, teacher can
add parts of his course to help students to understand it.

At second, for science purposes: while development of a new branch
of mathematics, scientists may include their work to the system to
avoid explanation while presenting his work to anybody.\\

Yet another artificial intelligence?

No. The system aims to be a DataBase. There is a wide range of
computing tools, automated theorem provers. WikiMathDB is not one of
them. WikiMathDB's responsibilities is storing and explaining, not
evaluation. There also is a huge number of computer-aided math
libraries, WikiMathDB differs from them by containing the logical
consistency, in the way how author thought.

\newpage

\section{Examples}

{\bf Example 1 (elementary)}.\\

\hrule\bigskip

{\bf Given}: $B \subset A, b \in B$, so $b \in A$.\\

\hrule\bigskip

Why "so"? Maybe in some theories it is not true? Here I can ask the
WikiMathDB to explain me why (by, for example, ctrl-clicking on the
word "so"). And get the answer based on definition of the subset
operator:
$$(B \subset A) \Leftrightarrow (\forall b \in B \Rightarrow b \in
A)$$ This answer is based on definition of \textit{subset}:
$$\forall X,Y - sets \Rightarrow [X \subset Y) \Leftrightarrow
(\forall x \in X \Rightarrow x \in Y)]$$ Where $B$ can be placed on
the place of $X$, $A$ --- on the place of $Y$, and $b$ --- on the
place of $x$.
\\

{\bf Example 2}.\\

\hrule\bigskip

{\bf Given}: \textbf{Basis} is a \textit{linearly
independent} \textit{spanning set}.\\

\hrule\bigskip

Having read this, one may want to know, what do they mean while
talking "linearly independent" and "spanning set"? By ctrl-clicking
on these terms he can get the definitions of the terms. Here
WikiMathDB can make an easy translation from English to math: "is\
a"$\ \rightarrow\; "\in"$, so, this definition will be in form:
$$
\forall b \in B \Rightarrow (b \in L \wedge b \in S)
$$
Where L is a set of linearly independent vector sets, and S --- of
spanning sets. Asking the system what does spanning set mean gives
the definition of the spanning set itself.

But, you should notice that here we omitted the field in which
"basis" is defining: subspace and vector space. Because "basis" will
be used if form "basis of the subspace M in vector space V".
Omitting these these details is not allowed in the WikiMathDB. The
details like that make possible to fully understand matter by
computer. So, writing a definition like for "basis", author should
consider these details while adding anything to the WikiMathDB.
\\
\newpage

{\bf Example 3 (more comprehensive)}.

Let's look at one solution in field of probability theory,
characteristic functions:\\

\hrule\bigskip

{\bf Given}: Random variable $\xi$ has the Couchy distribution
$$
p_{\xi}(x) = \frac{1}{\pi} \frac{a}{a^2 + x^2}
$$

{\bf Problem}: Find the characteristic function for $\xi$.

{\bf Solution}:
$$
f_{\xi}(t) = \frac{a}{\pi} \int_{-\infty}^\infty e^{itx}
\frac{dt}{x^2 + a^2} = \frac{a}{\pi} \int_{-\infty}^\infty
\frac{cos(t x)}{x^2 + a^2}dt
$$

Notice that $f_\xi(t)$ -- even function, so we only need to know its
values for $t>0$. Let us differentiate both sides of the last
equation, we'll get:

$$
f_{\xi}(t) = \frac{a}{\pi} \int_{-\infty}^\infty \frac{-x\ sin(t x)}
{x^2 + a^2}dt \eqno{(1)}
$$

It is known from the math analysis that

$$\int_{-\infty}^\infty \frac{sin(t x)}{x}dx = \pi
\parbox{4em}{\hspace{1em}, so\hfill} a= \frac{a}{\pi}\int_{-\infty}^\infty \frac{sin(t
x)}{x}dx\eqno{(2)}$$


Summarizing (1) and (2), we'll get

Let us differentiate both sides by $t$. We get $f''_\xi(t) = a^2
f_\xi(t)$. So, $f_\xi(t) = c_1 e^{at} + c_2 e^{-at}$, $t>0$. Because
of $f_\xi(t)$ -- bounded function on $\mathbb{R}$, then $c_1 = 0$,
and, basing on property $f_\xi(0) = 1$, get $c_2 = 1$. So, while $t
> 0 \; $ $\; f_\xi(t) = e^{-at}$, but taking consideration of the function
eveness, finally get $f_\xi(t) = e^{-a|t|}$, $t \in \mathbb{R}$.\\

\hrule\bigskip

Here we see a lot of hidden things: starting from definition of
characteristic function and ending with differentiation of improper
parametrized integral by parameter, along with on-the-fly solve of
Euler's differential equation. There are wasting a reader's time if
the reader wants to check (or understand) the solution. It is good
if reader is always a genius, but if he isn't? With the help of
WikiMathDB, author can waste his own time but save readers' time.
Also, explaining something in detail to the stupid computer can
eliminate all mistakes that human eye may not notice.

\newpage

\section{Global Design}

Version 1.0 should provide basic functionality for {\bf explaining}
and {\bf viewing} existent content, along with {\bf adding} new
content. These include:

\begin{enumerate}
\item
{\bf Common}. All development must use the portable technologies and
tools for compatibility, easier maintaining and development.
\item
{\bf Client side}. Basic client should be {\bf crossplatform}, has a
{\bf WYSIWYG math editor}. Added documents should be sent and stored
too, not only parsed and transformed versions. A web-application
form sounds to be a good idea.
\item
{\bf Server side}. Server side should provide an open, documented
API, to enable creating different clients for the same servers. The
SOAP technology sounds to be a good idea. Structure should consider
further features described below for future upgrades.
\end{enumerate}

Furter versions will point on:
\begin{itemize}
\item {\bf internationalization} -- distinguish different
languages, along with different terminologies for the same fields.
\item {\bf distributed server side} -- allow anyone to install their own
clones of the database, work with them, and send changes to the our
repository.
\item {\bf versioning} -- track changes made to the database,
allowing reverting changes made. Forking and merging will be a good
point too.
\item {\bf authorization} -- close some parts of database from public
edit.
\item {\bf heuristics} -- help editors to add new content by
performing evaluating and computing like it done by a lot of
computing environments like Mathematica\circledR.

\end{itemize}

\end{document}
